Implicit Holographic Memories

### 2016-11-07 papers/hrr-paper/paper.tex
Modified title are, abstract and introduction.

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Semantic Holographic Memories}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Igor Peric, Stefan Ulbrich, Alexandru Lesi, R{\"u}diger Dillman, Marius Z{\"o}llner}
\IEEEauthorblockA{FZI Forschungszentrum Informatik\\
Karlsruhe, Germany\\
Email: $\{$peric, ulbrich, dillmann, zoellner$\}$@fzi.de}
}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Vector Symbolic Architectures (VSA) define a set of operations for storage (compression), manipulation and retrieval of symbolic concepts, represented as fixed-length vectors. A specific instance of VSA, Holographic Reduced Representations (HRR), has proven to exhibit properties similar to human short-term memory and is, as such, interesting for computational modeling.

In this paper we extend HRR approach by introducing topology-preserving, invertible encoding/decoding procedure for all major data types used in robotics and engineering. These encoding schemes play the role of otherwise used clean-up modules after memory retrieval (e.g. self-organizing maps), thus eliminating necessity for them.

We test our memories in three different robotic tasks: mapping and recall, visual object classification and state machine scripting (holographic controllers).
\end{abstract}

@@

Vector Symbolic Architectures (VSA) define a set of operations for storage (compression) and retrieval of symbolic concepts in fixed size vectors. In its basic form every VSA has to implement three basic operations:
\begin{enumerate}
\item association (binding) $\mapsto \otimes$
\item superposition (appending) $\mapsto +$
\item probing (unbinding) $\mapsto \oslash$
\end{enumerate}

In addition, every VSA can be considered as a compression technique, since binary operations on N-element vectors return N-element vectors as well.

\textbf{Association} (also called "binding") is operation of linking two operands together, resulting in a vector of same size which is dissimilar (orthogonal) to both original vectors, but contains scrambled information from both of the operands. For instance, linking property "red" to object "triangle" can be done using binding operation:

@@

Process of converting noisy result of probing into the closest know symbol is called \textbf{clean-up procedure}. Straightforward way of implementing clean-up memory is using lookup table, while other researchers show that any type of attractor network can be used (e.g. self-organizing maps).

\section{Holographic reduced representations}

@@

\begin{figure}
\begin{subfigure}{1\columnwidth}
\includegraphics[width=\columnwidth]{img/scalar-pre-perm.png}
\caption{Binging}
\label{no-perm-a}
\end{subfigure}
\begin{subfigure}{1\columnwidth}
\includegraphics[width=\columnwidth]{img/scalar-pre-perm-probe.png}
\caption{Probing}
\label{no-perm-b}
\end{subfigure}
\caption{Plot of vector values. Figure \ref{no-perm-a} shows the binding of the value 8 on a scale to 10 with the symbolic representation of "ball", which is random. In figure \ref{no-perm-b} "ball" is being probed, and the expected result is a Gaussian bell for 8, mixed with some noise.}
\label{no-perm}
\end{figure}

@@

\begin{figure}
\begin{subfigure}{1\columnwidth}
\includegraphics[width=\columnwidth]{img/scalar-post-perm.png}
\caption{Binging}
\label{perm-a}
\end{subfigure}
\begin{subfigure}{1\columnwidth}
\includegraphics[width=\columnwidth]{img/scalar-post-perm-probe.png}
\caption{Probing}
\label{perm-b}
\end{subfigure}
\caption{Figure \ref{perm-a} shows the binding of the permuted Gaussian for value 8 on a scale to 10 with the symbolic representation of "ball", which is random. In figure \ref{perm-b} "ball" is being probed, and the expected result is a Gaussian bell for 8, mixed with some noise.}
\label{perm}
\end{figure}

@@

\section{Conclusions}

@@

\bibliographystyle{IEEEtran}

\bibliography{IEEEabrv,bibliography}

>>>>

\usepackage{cite}

@@

\begin{document}
	%
	% paper title
	% Titles are generally capitalized except for words such as a, an, and, as,
	% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
	% not capitalized unless they are the first or last word of the title.
	% Linebreaks \\ can be used within to get better formatting as desired.
	% Do not put math or special symbols in the title.
	\title{Implicit Holographic Memories}

	% author names and affiliations
	% use a multiple column layout for up to three different
	% affiliations
	\author{\IEEEauthorblockN{Igor Peric, Stefan Ulbrich, Alexandru Lesi, Daniel Spies, R{\"u}diger Dillman, Marius Z{\"o}llner}
		\IEEEauthorblockA{FZI Forschungszentrum Informatik\\
			Karlsruhe, Germany\\
			Email: $\{$peric, ulbrich, lesi, spies, dillmann, zoellner$\}$@fzi.de}
	}

@@

	% As a general rule, do not put math, special symbols or citations
	% in the abstract
	\begin{abstract}
		Vector Symbolic Architectures (VSA) define a set of operations for storage (compression), manipulation and retrieval of symbolic concepts, represented as fixed-length vectors from ${\rm I\!R}^N$. A specific instance of VSA, Holographic Reduced Representations (HRR), has proven to exhibit properties similar to human short-term memory and is, as such, interesting for computational modeling.

		In this paper we extend HRR approach by introducing implicit, topology-preserving, encoding/decoding procedure using radial basis functions. Extended approach involves randomly chosen permutation (for encoding) and its inverse (for decoding), necessary for uniform distribution of signal in Fourier space in which embedding takes place. These novel encoding schemes, together with extended operation pipeline, play the role of so called clean-up modules after memory retrieval (e.g. self-organizing maps). Effectively, encoding implicitly represents the symbol, so no further lookup is needed.

		We show that newly introduced encoding scheme didn't have negative impact on memory capacity in comparison with original author's capacity benchmark. We also test our memories in three different robotics tasks: mapping and recall of an area by a mobile agent, top-down visual attention shift and state machine scripting (holographic controllers).
	\end{abstract}

@@

	Vector Symbolic Architectures (VSA) define a set of operations for storage (compression) and retrieval of symbolic concepts in fixed size vectors. As described in \cite{vsa1}, every VSA has to implement three basic operations:
	\begin{enumerate}
		\item association (binding) $\mapsto \otimes$
		\item superposition (appending) $\mapsto +$
		\item probing (unbinding) $\mapsto \oslash$
	\end{enumerate}

	In addition, every VSA can be considered as a lossy compression technique, since binary operations on N-element vectors return N-element vectors as well.

	Depending of the way these three operations are implemented we can distinguish several popular VSAs, such as Holographic Reduced Representations \cite{hrr-1}, Circular HRRs \cite{circular-hrr}, Binary Spatter Codes \cite{bsc-1}, Holographic Graph Neurons\cite{holographic-graph-nouron}, Geometric Algebras\cite{geometric-algebra} and Random Permutations. The most widely used ones are HRRs and they are briefly explained in section \ref{sec:hrr}. Authors in \cite{learning-language} have used them to solve word embedding problem for extremely large dataset by replacing computationally expensive big matrix factorization with iterative estimation algorithm based on HRRs. Some of the other work is centered around pattern recognition \cite{pattern-recognition}, functional large-scale cognitive architecture models \cite{spaun} and representing scene structure \cite{scene-structure}. Some preliminary theoretical assumptions have been made about ability to use them as agent controllers\cite{embedding-behaviour}, where a symbolic program is represented with a set of combined symbolic actions associated with actuator symbols and symbolic sensing.

	Authors in \cite{spaun} have demonstrated functional cognitive model Spaun, implemented as HRRs using biologically plausible models of neuron dynamics, is capable of performing 7 different cognitive tasks. In serial memory task they show that model of forgetting exhibits similar behavior as human working memory due to limited capacity. Besides that, they report similarity between human working memory capacity and average number of items (symbols) that their model is capable of holding in memory. This successful model of forgetting is one of the reasons our work is centered around HRRs.

	Second reason is the fact that HRRs are the only VSAs up to date operating on convolution operations in their core. This is appealing if we take into consideration numerous recent advancement in deep learning and convolutional neural networks. There are two basic reasons for success of these networks: training on big data and using convolution as the main architectural constraint. Although description of biophysical mechanism underlying convolution and concept of weight sharing is still missing, it is important to reuse already proven concepts (such as convolutional filtering) in as general way as possible. In other words, addressing multiple fundamentally different problems with the same or similar computational principles can be considered as a step towards unified brain processing theory.

	\section{Vector Symbolic Architectures}

	\textbf{Association} (also called "binding") is an operation of linking two operands together, resulting in a vector of the same size which is dissimilar (orthogonal) to both original vectors, but contains scrambled information from both of the operands. For instance, linking the property "red" to object "triangle" can be done using binding operation:

@@

	Process of converting noisy result of probing into the closest know symbol is called \textbf{clean-up procedure}. Straightforward way of implementing clean-up memory is using a lookup table, while other researchers show that any type of attractor network can be used (e.g. self-organizing maps).

	\subsection{Holographic reduced representations}
	\label{sec:hrr}

@@

	\begin{figure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-pre-perm.png}
			\caption{Binding}
			\label{no-perm-a}
		\end{subfigure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-pre-perm-probe.png}
			\caption{Probing}
			\label{no-perm-b}
		\end{subfigure}
		\caption{Plot of vector values. Figure \ref{no-perm-a} shows the binding of the value 8 on a scale to 10 with the symbolic representation of "ball", which is random. In figure \ref{no-perm-b} "ball" is being probed, and the expected result is a Gaussian bell for 8, mixed with some noise.}
		\label{no-perm}
	\end{figure}

@@

	\begin{figure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-post-perm.png}
			\caption{Binding}
			\label{perm-a}
		\end{subfigure}
		\begin{subfigure}{1\columnwidth}
			\includegraphics[width=\columnwidth]{img/scalar-post-perm-probe.png}
			\caption{Probing}
			\label{perm-b}
		\end{subfigure}
		\caption{Figure \ref{perm-a} shows the binding of the permuted Gaussian for value 8 on a scale to 10 with the symbolic representation of "ball", which is random. In figure \ref{perm-b} "ball" is being probed, and the expected result is a Gaussian bell for 8, mixed with some noise.}
		\label{perm}
	\end{figure}

@@

	\section{Conclusions}
	We have introduced a mechanism for intrinsic symbol encoding/decoding for Holographic Reduced Representations. As a consequence, our approach eliminates the need for clean-up memories as they are described in literature. Instead, result of probing operation directly gives interpretable symbol, which can be decoded in constant computational time. In other words, time needed for encoding/decoding is completely ignorant of total number of symbols memorized, allowing memories that can store large corpus of different symbols.
	A concrete biophysical evidence and theoretical model of implementation of convolution in brain is, unfortunately, still missing.

@@

	\include{bibliography}
